\begin{thebibliography}{63}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Bergero et~al.(2007)Bergero, Forrest, Kamau, and
  Charlesworth]{Bergero:2007kr}
R~Bergero, A~Forrest, E~Kamau, and D~Charlesworth.
\newblock {Evolutionary strata on the X chromosomes of the dioecious plant
  \emph{Silene latifolia}: evidence from new sex-linked genes}.
\newblock \emph{Genetics}, 175\penalty0 (4):\penalty0 1945--1954, February
  2007.

\bibitem[Nam and Ellegren(2008)]{Nam:2008ko}
K~Nam and H~Ellegren.
\newblock {The chicken (\emph{Gallus gallus}) Z chromosome contains at least
  three nonlinear evolutionary strata}.
\newblock \emph{Genetics}, 180\penalty0 (2):\penalty0 1131--1136, September
  2008.

\bibitem[Lemaitre et~al.(2009)Lemaitre, Braga, Gautier, Sagot, Tannier, and
  Marais]{Lemaitre:2009ij}
C~Lemaitre, M~D~V Braga, C~Gautier, M~F Sagot, E~Tannier, and G~A~B Marais.
\newblock {Footprints of inversions at present and past pseudoautosomal
  boundaries in human sex chromosomes}.
\newblock \emph{Genome Biology and Evolution}, 1\penalty0 (0):\penalty0 56--66,
  April 2009.

\bibitem[Wang et~al.(2012)Wang, Na, Yu, Gschwend, Han, Zeng, Aryal, VanBuren,
  Murray, Zhang, Navajas-P{\'e}rez, Feltus, Lemke, Tong, Chen, Wai, Singh,
  Wang, Min, Alam, Charlesworth, Moore, Jiang, Paterson, and Ming]{Wang:2012}
Jianping Wang, Jong-Kuk Na, Qingyi Yu, Andrea~R Gschwend, Jennifer Han,
  Fanchang Zeng, Rishi Aryal, Robert VanBuren, Jan~E Murray, Wenli Zhang,
  Rafael Navajas-P{\'e}rez, F~Alex Feltus, Cornelia Lemke, Eric~J Tong, Cuixia
  Chen, Ching~Man Wai, Ratnesh Singh, Ming-Li Wang, Xiang~Jia Min, Maqsudul
  Alam, Deborah Charlesworth, Paul~H Moore, Jiming Jiang, Andrew~H Paterson,
  and Ray Ming.
\newblock {Sequencing papaya X and Yh chromosomes reveals molecular basis of
  incipient sex chromosome evolution.}
\newblock \emph{Proceedings of the National Academy of Sciences}, 109\penalty0
  (34):\penalty0 13710--13715, August 2012.

\bibitem[Charlesworth(2013)]{Charlesworth:2013}
Deborah Charlesworth.
\newblock {Plant sex chromosome evolution.}
\newblock \emph{Journal of Experimental Botany}, 64\penalty0 (2):\penalty0
  405--420, January 2013.

\bibitem[Rice(1996)]{Rice:1996ke}
William~R Rice.
\newblock {Evolution of the Y Sex Chromosome in Animals}.
\newblock \emph{BioScience}, 46\penalty0 (5):\penalty0 331--343, May 1996.

\bibitem[Charlesworth and Charlesworth(2000)]{Charlesworth:2000cc}
B~Charlesworth and D~Charlesworth.
\newblock {The degeneration of Y chromosomes.}
\newblock \emph{Philosophical transactions of the Royal Society of London.
  Series B, Biological sciences}, 355\penalty0 (1403):\penalty0 1563--1572,
  November 2000.

\bibitem[Bachtrog(2006)]{Bachtrog:2006ed}
Doris Bachtrog.
\newblock {A dynamic view of sex chromosome evolution.}
\newblock \emph{Current opinion in genetics {\&} development}, 16\penalty0
  (6):\penalty0 578--585, December 2006.

\bibitem[Marais et~al.(2008)Marais, Nicolas, Bergero, Chambrier, Kejnovsky,
  Mon{\'e}ger, Hobza, Widmer, and Charlesworth]{Marais:2008hm}
Gabriel A~B Marais, Michael Nicolas, Roberta Bergero, Pierre Chambrier, Eduard
  Kejnovsky, Fran{\c c}oise Mon{\'e}ger, Roman Hobza, Alex Widmer, and Deborah
  Charlesworth.
\newblock {Evidence for degeneration of the Y chromosome in the dioecious plant
  \emph{Silene latifolia}}.
\newblock \emph{Current Biology}, 18\penalty0 (7):\penalty0 545--549, April
  2008.

\bibitem[Charlesworth et~al.(2005)Charlesworth, Charlesworth, and
  Marais]{Charlesworth:2005cx}
D~Charlesworth, B~Charlesworth, and G~Marais.
\newblock {Steps in the evolution of heteromorphic sex chromosomes}.
\newblock \emph{Heredity}, 95\penalty0 (2):\penalty0 118--128, May 2005.

\bibitem[Charlesworth(2015)]{Charlesworth:2015dj}
Deborah Charlesworth.
\newblock {Plant contributions to our understanding of sex chromosome
  evolution}.
\newblock \emph{The New phytologist}, 208\penalty0 (1):\penalty0 52--65, 2015.

\bibitem[Nei(1969)]{Nei:1969tb}
Masatoshi Nei.
\newblock {Linkage modification and sex difference in recombination}.
\newblock \emph{Genetics}, 63\penalty0 (3):\penalty0 681--699, November 1969.

\bibitem[Charlesworth and Charlesworth(1978)]{Charlesworth:1978ut}
B~Charlesworth and D~Charlesworth.
\newblock {A model for the evolution of dioecy and gynodioecy}.
\newblock \emph{American Naturalist}, 122\penalty0 (988):\penalty0 975--997,
  1978.

\bibitem[Fisher(1931)]{Fisher:1931jo}
R~Fisher.
\newblock {The evolution of dominance}.
\newblock \emph{Biological Reviews}, 6\penalty0 (4):\penalty0 363, October
  1931.

\bibitem[Bull(1983)]{Bull:1983vi}
J~J Bull.
\newblock \emph{{Evolution of sex determining mechanisms.}}
\newblock The Benjamin/Cummings Publishing Company, Inc., 1983.

\bibitem[Rice(1987)]{Rice:1987hs}
William~R Rice.
\newblock {The accumulation of sexually antagonistic genes as a selective agent
  promoting the evolution of reduced recombination between primitive sex
  chromosomes}.
\newblock \emph{Evolution}, 41\penalty0 (4):\penalty0 911, July 1987.

\bibitem[Charlesworth and Charlesworth(1980)]{Charlesworth:1980}
D~Charlesworth and B~Charlesworth.
\newblock {Sex differences in fitness and selection for centric fusions between
  sex-chromosomes and autosomes}.
\newblock \emph{Genetical Research}, 35\penalty0 (02):\penalty0 205--214, 1980.

\bibitem[Lenormand(2003)]{Lenormand:2003ug}
Thomas Lenormand.
\newblock {The evolution of sex dimorphism in recombination.}
\newblock \emph{Genetics}, 163\penalty0 (2):\penalty0 811--822, 2003.

\bibitem[Otto et~al.(2011)Otto, Pannell, Peichel, and Ashman]{Otto:2011ua}
S~P Otto, J~R Pannell, C~L Peichel, and T-L Ashman.
\newblock {About PAR: the distinct evolutionary dynamics of the pseudoautosomal
  region}.
\newblock \emph{Trends in Genetics}, 27\penalty0 (9):\penalty0 358--367, 2011.

\bibitem[Mulcahy et~al.(1996)Mulcahy, Sari-Gorla, and Mulcahy]{Mulcahy:1996ha}
David~L Mulcahy, Mirella Sari-Gorla, and Gabriella~Bergamini Mulcahy.
\newblock {Pollen selection - past, present and future}.
\newblock \emph{Sexual Plant Reproduction}, 9\penalty0 (6):\penalty0 353--356,
  November 1996.

\bibitem[Bernasconi(2004)]{Bernasconi:2004dk}
G~Bernasconi.
\newblock {Evolutionary ecology of the prezygotic stage}.
\newblock \emph{Science}, 303\penalty0 (5660):\penalty0 971--975, February
  2004.

\bibitem[Joseph and Kirkpatrick(2004)]{JOSEPH:2004haa}
S~Joseph and Mark Kirkpatrick.
\newblock {Haploid selection in animals}.
\newblock \emph{Trends in Ecology {\&} Evolution}, 19\penalty0 (11):\penalty0
  592--597, November 2004.

\bibitem[Immler et~al.(2012)Immler, Arnqvist, and Otto]{Immler:2012tl}
S~Immler, G~Arnqvist, and S~P Otto.
\newblock {Ploidally antagonistic selection maintains stable genetic
  polymorphism}.
\newblock \emph{Evolution}, 66\penalty0 (1):\penalty0 55--65, 2012.

\bibitem[Skogsmyr and Lankinen(2002)]{SKOGSMYR:2002ce}
I~Skogsmyr and A~Lankinen.
\newblock {Sexual selection: an evolutionary force in plants?}
\newblock \emph{Biological Reviews}, 77\penalty0 (4):\penalty0 537--562,
  November 2002.

\bibitem[Moore and Pannell(2011)]{Moore:2011jt}
Jamie~C Moore and John~R Pannell.
\newblock {Sexual selection in plants}.
\newblock \emph{Current Biology}, 21\penalty0 (5):\penalty0 R176--R182, March
  2011.

\bibitem[Marshall and Evans(2016)]{Marshall:2016fe}
D~L Marshall and A~S Evans.
\newblock {Can selection on a male mating character result in evolutionary
  change? A selection experiment on California wild radish, \emph{Raphanus
  sativus}}.
\newblock \emph{American journal of botany}, 103\penalty0 (3):\penalty0
  553--567, March 2016.

\bibitem[Borg et~al.(2009)Borg, Brownfield, and Twell]{Borg:2009jpa}
M~Borg, L~Brownfield, and D~Twell.
\newblock {Male gametophyte development: a molecular perspective}.
\newblock \emph{Journal of Experimental Botany}, 60\penalty0 (5):\penalty0
  1465--1478, January 2009.

\bibitem[Arunkumar et~al.(2013)Arunkumar, Josephs, Williamson, and
  Wright]{Arunkumar:2013iq}
R~Arunkumar, E~B Josephs, R~J Williamson, and S~I Wright.
\newblock {Pollen-specific, but not sperm-specific, genes show stronger
  purifying selection and higher rates of positive selection than sporophytic
  genes in \emph{Capsella grandiflora}}.
\newblock \emph{Molecular biology and evolution}, 30\penalty0 (11):\penalty0
  2475--2486, October 2013.

\bibitem[Gossmann et~al.(2014)Gossmann, Schmid, Grossniklaus, and
  Schmid]{Gossmann:2014dua}
T~I Gossmann, M~W Schmid, U~Grossniklaus, and K~J Schmid.
\newblock {Selection-driven evolution of sex-biased genes Is consistent with
  sexual selection in \emph{Arabidopsis thaliana}}.
\newblock \emph{Molecular biology and evolution}, 31\penalty0 (3):\penalty0
  574--583, February 2014.

\bibitem[Hedhly et~al.(2004)Hedhly, Hormaza, and Herrero]{Hedhly:2004iv}
A~Hedhly, J~I Hormaza, and M~Herrero.
\newblock {Effect of temperature on pollen tube kinetics and dynamics in sweet
  cherry, \emph{Prunus avium} (Rosaceae).}
\newblock \emph{American journal of botany}, 91\penalty0 (4):\penalty0
  558--564, April 2004.

\bibitem[Clarke et~al.(2004)Clarke, Khan, and Siddique]{Clarke:2004ir}
Heather~J Clarke, Tanveer~N Khan, and Kadambot H~M Siddique.
\newblock {Pollen selection for chilling tolerance at hybridisation leads to
  improved chickpea cultivars}.
\newblock \emph{Euphytica}, 139\penalty0 (1):\penalty0 65--74, 2004.

\bibitem[Frascaroli and Songstad(2001)]{Frascaroli:2001ee}
E~Frascaroli and D~D Songstad.
\newblock {Pollen genotype selection for a simply inherited qualitative factor
  determining resistance to chlorsulfuron in maize}.
\newblock \emph{Theoretical and Applied Genetics}, 102\penalty0 (2-3):\penalty0
  342--346, February 2001.

\bibitem[Searcy and Mulcahy(1985)]{Searcy:1985vt}
K~B Searcy and D~L Mulcahy.
\newblock {Pollen selection and the gametophytic expression of metal tolerance
  in \emph{Silene dioica} (Caryophyllaceae) and \emph{Mimulus guttatus}
  (Scrophulariaceae)}.
\newblock \emph{American journal of botany}, 72\penalty0 (11):\penalty0
  1700--1706, 1985.

\bibitem[Ravikumar et~al.(2003)Ravikumar, Patil, and
  Salimath]{Ravikumar:2003uo}
R~L Ravikumar, B~S Patil, and P~M Salimath.
\newblock {Drought tolerance in sorghum by pollen selection using osmotic
  stress}.
\newblock \emph{Euphytica}, 133:\penalty0 371--376, 2003.

\bibitem[Ravikumar et~al.(2012)Ravikumar, Chaitra, Choukimath, and
  Soregaon]{Ravikumar:2012ej}
R~L Ravikumar, G~N Chaitra, Anilkumar~M Choukimath, and C~D Soregaon.
\newblock {Gametophytic selection for wilt resistance and its impact on the
  segregation of wilt resistance alleles in chickpea (\emph{Cicer arietinum}
  L.)}.
\newblock \emph{Euphytica}, 189\penalty0 (2):\penalty0 173--181, July 2012.

\bibitem[Hecht(1998)]{Hecht:1998hz}
Norman~B Hecht.
\newblock {Molecular mechanisms of male germ cell differentiation}.
\newblock \emph{Bioessays}, 20\penalty0 (7):\penalty0 555--561, December 1998.

\bibitem[Zheng et~al.(2001)Zheng, Deng, and Martin-DeLeon]{Zheng:2001fi}
Y~Zheng, X~Deng, and P~A Martin-DeLeon.
\newblock {Lack of sharing of Spam1 (Ph-20) among mouse spermatids and
  transmission ratio distortion}.
\newblock \emph{Biology of Reproduction}, 64\penalty0 (6):\penalty0 1730--1738,
  June 2001.

\bibitem[Vibranovski et~al.(2010)Vibranovski, Chalopin, Lopes, Long, and
  Karr]{Vibranovski:2010et}
M~D Vibranovski, D~S Chalopin, H~F Lopes, M~Long, and T~L Karr.
\newblock {Direct evidence for postmeiotic transcription during
  \emph{Drosophila melanogaster} spermatogenesis}.
\newblock \emph{Genetics}, 186\penalty0 (1):\penalty0 431--433, September 2010.

\bibitem[Immler et~al.(2014)Immler, Hotzy, Alavioon, Petersson, and
  Arnqvist]{Immler:2014im}
S~Immler, C~Hotzy, G~Alavioon, E~Petersson, and G~Arnqvist.
\newblock {Sperm variation within a single ejaculate affects offspring
  development in Atlantic salmon}.
\newblock \emph{Biology letters}, 10\penalty0 (2):\penalty0 20131040, February
  2014.

\bibitem[Otto and Day(2007)]{Otto:2007tv}
S~P Otto and T~Day.
\newblock \emph{{A biologist's guide to mathematical modeling in ecology and
  evolution}}.
\newblock Princeton University Pres, Princeton, NJ, 2007.

\bibitem[Otto(2014)]{Otto:2014jf}
S~P Otto.
\newblock {Selective maintenance of recombination between the sex chromosomes.}
\newblock \emph{Journal of Evolutionary Biology}, 27\penalty0 (7):\penalty0
  1431--1442, July 2014.

\bibitem[Charlesworth and Wall(1999)]{Charlesworth:1999kd}
B~Charlesworth and J~D Wall.
\newblock {Inbreeding, heterozygote advantage and the evolution of neo-X and
  neo-Y sex chromosomes}.
\newblock \emph{Proceedings of the Royal Society B: Biological Sciences},
  266\penalty0 (1414):\penalty0 51--56, January 1999.

\bibitem[Pennell et~al.(2015)Pennell, Kirkpatrick, Otto, Vamosi, Peichel,
  Valenzuela, and Kitano]{Pennell:2015im}
Matthew~W Pennell, Mark Kirkpatrick, Sarah~P Otto, Jana~C Vamosi, Catherine~L
  Peichel, Nicole Valenzuela, and Jun Kitano.
\newblock {Y fuse? Sex chromosome fusions in fishes and reptiles}.
\newblock \emph{PLOS Genetics}, 11\penalty0 (5):\penalty0 e1005237, May 2015.

\bibitem[Karlin and McGregor(1972{\natexlab{a}})]{Karlin:1972ab}
Samuel Karlin and James McGregor.
\newblock {Application of method of small parameters to multi-niche population
  genetic models}.
\newblock \emph{Theoretical Population Biology}, 3\penalty0 (2):\penalty0
  186--209, June 1972{\natexlab{a}}.

\bibitem[Karlin and McGregor(1972{\natexlab{b}})]{Karlin:1972dq}
Samuel Karlin and James McGregor.
\newblock {Polymorphisms for genetic and ecological systems with weak
  coupling}.
\newblock \emph{Theoretical Population Biology}, 3\penalty0 (2):\penalty0
  210--238, June 1972{\natexlab{b}}.

\bibitem[Field et~al.(2013)Field, Pickup, and Barrett]{Field:2013cc}
David~L Field, Melinda Pickup, and Spencer C~H Barrett.
\newblock {Comparative analyses of sex-ratio variation in dioecious flowering
  plants.}
\newblock \emph{Evolution}, 67\penalty0 (3):\penalty0 661--672, March 2013.

\bibitem[Conn and Blum(1981)]{Conn:1981uw}
J~S Conn and Udo Blum.
\newblock {Sex ratio of \emph{Rumex hastatulus}: the effect of environmental
  factors and certation}.
\newblock \emph{Evolution}, 35\penalty0 (6):\penalty0 1108--1116, 1981.

\bibitem[Stehlik and Barrett(2006)]{Stehlik:2006to}
Ivana Stehlik and Spencer C~H Barrett.
\newblock {Pollination intensity influences sex ratios in dioecious Rumex
  nivalis, a wind-pollinated plant.}
\newblock \emph{Evolution}, 60\penalty0 (6):\penalty0 1207--1214, June 2006.

\bibitem[Field et~al.(2012)Field, Pickup, and Barrett]{Field:2012fd}
David~L Field, Melinda Pickup, and Spencer C~H Barrett.
\newblock {The influence of pollination intensity on fertilization success,
  progeny sex ratio, and fitness in a wind-pollinated, dioecious plant}.
\newblock \emph{International Journal of Plant Sciences}, 173\penalty0
  (2):\penalty0 184--191, February 2012.

\bibitem[Lloyd(1974)]{Lloyd:1974tz}
D~G Lloyd.
\newblock \emph{{Female-predominant sex ratios in angiosperms}}, volume~32.
\newblock Heredity, 1974.

\bibitem[Stehlik and Barrett(2005)]{Stehlik:2005ul}
I~Stehlik and SCH Barrett.
\newblock {Mechanisms governing sex-ratio variation in dioecious \emph{Rumex
  nivalis}}.
\newblock \emph{Evolution}, 59\penalty0 (4):\penalty0 814--825, 2005.

\bibitem[Ming et~al.(2011)Ming, Bendahmane, and Renner]{Ming:2011iy}
Ray Ming, Abdelhafid Bendahmane, and Susanne~S Renner.
\newblock {Sex chromosomes in land plants}.
\newblock \emph{dx.doi.org}, 62\penalty0 (1):\penalty0 485--514, June 2011.

\bibitem[Bachtrog et~al.(2014)Bachtrog, Mank, Peichel, Kirkpatrick, Otto,
  Ashman, Hahn, Kitano, Mayrose, Ming, Perrin, Ross, Valenzuela, Vamosi, and
  {Tree of Sex Consortium}]{Bachtrog:2014bx}
Doris Bachtrog, Judith~E Mank, Catherine~L Peichel, Mark Kirkpatrick, Sarah~P
  Otto, Tia-Lynn Ashman, Matthew~W Hahn, Jun Kitano, Itay Mayrose, Ray Ming,
  Nicolas Perrin, Laura Ross, Nicole Valenzuela, Jana~C Vamosi, and {Tree of
  Sex Consortium}.
\newblock {Sex determination: why so many ways of doing it?}
\newblock \emph{PLoS Biol}, 12\penalty0 (7):\penalty0 e1001899, July 2014.

\bibitem[Vicoso and Bachtrog(2015)]{Vicoso:2015hf}
Beatriz Vicoso and Doris Bachtrog.
\newblock {Numerous transitions of sex chromosomes in Diptera.}
\newblock \emph{PLoS Biol}, 13\penalty0 (4):\penalty0 e1002078, April 2015.

\bibitem[Travers and Mazer(2001)]{Travers:2001vx}
S~E Travers and S~J Mazer.
\newblock {Trade-offs between male and female reproduction associated with
  allozyme variation in phosphoglucoisomerase in an annual plant (\emph{Clarkia
  unguiculata}: Onagraceae).}
\newblock \emph{Evolution}, 55\penalty0 (12):\penalty0 2421--2428, December
  2001.

\bibitem[Lenormand and Dutheil(2005)]{Lenormand:2005vb}
T~Lenormand and J~Dutheil.
\newblock {Recombination difference between sexes: a role for haploid
  selection}.
\newblock \emph{PLoS Biol}, 3\penalty0 (3):\penalty0 e63, 2005.

\bibitem[van Doorn and Kirkpatrick(2007)]{vanDoorn:2007eu}
G~S van Doorn and Mark Kirkpatrick.
\newblock {Turnover of sex chromosomes induced by sexual conflict.}
\newblock \emph{Nature}, 449\penalty0 (7164):\penalty0 909--912, October 2007.

\bibitem[Feldman and Otto(1989)]{Feldman:1989gm}
Marcus~W Feldman and Sarah~P Otto.
\newblock {More on recombination and selection in the modifier theory of
  sex-ratio distortion}.
\newblock \emph{Theoretical Population Biology}, 35\penalty0 (2):\penalty0
  207--225, April 1989.

\bibitem[Haig(2010)]{Haig:2010em}
David Haig.
\newblock {Games in tetrads: segregation, recombination, and meiotic drive.}
\newblock \emph{The American Naturalist}, 176\penalty0 (4):\penalty0 404--413,
  October 2010.

\bibitem[Brandvain and Coop(2012)]{Brandvain:2012hh}
Y~Brandvain and G~Coop.
\newblock {Scrambling eggs: meiotic drive and the evolution of female
  recombination rates}.
\newblock \emph{Genetics}, 190\penalty0 (2):\penalty0 709--723, February 2012.

\bibitem[Patten(2014)]{Patten:2014tn}
M~M Patten.
\newblock {Meiotic drive influences the outcome of sexually antagonistic
  selection at a linked locus}.
\newblock \emph{Journal of Evolutionary Biology}, 27:\penalty0 2360--2370,
  2014.

\bibitem[Rydzewski et~al.(2016)Rydzewski, Carioscia, Li{\'e}vano, Lynch, and
  Patten]{Rydzewski:2016csa}
William~T Rydzewski, Sara~A Carioscia, Gabriela Li{\'e}vano, Victoria~D Lynch,
  and Manus~M Patten.
\newblock {Sexual antagonism and meiotic drive cause stable linkage
  disequilibrium and favour reduced recombination on the X chromosome}.
\newblock \emph{Journal of Evolutionary Biology}, 29\penalty0 (6):\penalty0
  1247--1256, March 2016.

\bibitem[{\'U}beda et~al.(2015){\'U}beda, Patten, and Wild]{Ubeda:2015fx}
Francisco {\'U}beda, Manus~M Patten, and Geoff Wild.
\newblock {On the origin of sex chromosomes from meiotic drive.}
\newblock \emph{Proceedings of the Royal Society B: Biological Sciences},
  282\penalty0 (1798):\penalty0 20141932, January 2015.

\end{thebibliography}
